CrypTool Script\footnote{ Background reading for the CrypTool program}\\ Mathematics and Cryptography

### 2003-05-30 script/de/chapters/numbertheory.tex
fixes for beta6

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Gruppen und modulare Arithmetik "uber $\mathbb{Z}_n $ und $\mathbb{Z}^*_n $}
\index{Gruppen}
In der Zahlentheorie und in der Kryptographie spielen mathematische 
\glqq {\em Gruppen}'' eine entscheidende Rolle. Von Gruppen spricht man nur, wenn f"ur
eine definierte Menge und eine definierte Relation (eine Operation wie
Addition oder Multiplikation) die folgenden Eigenschaften erf"ullt sind:
\begin{itemize}
\item Abgeschlossenheit \index{Abgeschlossenheit}
\item Existenz des neutralen Elements
\item Existenz des inversen Elements und
\item G"ultigkeit des Assoziativgesetzes.
\end{itemize}
Die abgek"urzte mathematische Schreibweise lautet: $(G, +)$ oder $(G,*)$.
\begin{definition}\label{def-zth-zn}
$\mathbb{Z}_n$ :\index{Z@$\mathbb{Z}_n$}
$$\mathbb{Z}_n \text{ umfasst alle ganzen Zahlen von } 0 \text{ bis } n-1: \mathbb{Z}_n = \{0, 1, 2,\cdots, n-2, n-1\}.$$
\end{definition}
$\mathbb{Z}_n$ ist eine h"aufig verwendete endliche Gruppe aus den nat"urlichen Zahlen. Sie 
wird manchmal auch als Restemenge $R$ modulo $n$ bezeichnet.

@@

{\bf Beispiel 2:} \\
Die folgende Tabelle\footnote{%
In \hyperlink{AppArith3b}{Anhang D zu diesem Kapitel} Finden Sie den Quelltext zur Berechnung der 
Tabelle mit Mathematica und Pari-GP.
} zeigt die Werte $a^i{\rm ~mod~}45$
f"ur die Exponenten $i = 1, 2, \cdots, 12$ und f"ur die Basen $a = 1, 2, \cdots, 12$
sowie den sich f"ur jedes $a$ daraus ergebenden Wert $ord_{45}(a)$.

@@

\newpage
{\bf Beispiel 4:} \\
Die folgende Tabelle\footnote{%
In \hyperlink{AppArith3c}{Anhang D zu diesem Kapitel} Finden Sie den Quelltext zur Berechnung der Tabelle mit Mathematica und Pari-GP.
} beantwortet die Frage, ob die Zahl $7$ eine Primitivwurzel von $46$ ist.
Die Voraussetzung/Bedingung ggT$(7,46)=1$ ist erf"ullt.

@@

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Beweis des RSA-Verfahrens mit Euler-Fermat}\label{rsabeweis}
\hypertarget{RSABeweis}{} Mit dem Satz von Euler-Fermat kann man in der Gruppe $\mathbb{Z}_n^*$ das 
RSA\footnote{%
Das RSA-Verfahren\index{RSA} ist das verbreitetste asymmetrische \index{Verschl""usselung!asymmetrisch} Kryptoverfahren. Es
wurde 1978 von Ronald Rivest, Adi Shamir und Leonard Adleman entwickelt und
kann sowohl zum Signieren wie zum Verschl"usseln eingesetzt werden.
Kryptographen assoziieren mit der Abk"urzung \glqq RSA'' immer dieses
Verfahren $-$ die folgende Anmerkung soll eher humorvoll zeigen, dass man jede
Buchstabenkombination mehrfach belegen kann: in Deutschland gibt es sehr
Interessen-lastige Diskussionen im Gesundheitswesen. Dabei wird mit \glqq RSA''
der vom Gesundheitsministerium geregelte \glqq {\bf R}isiko{\bf S}truktur{\bf A}usgleich'' in der
gesetzlichen Krankenversicherung mit einem j"ahrlichen Volumen von "uber 24
Milliarden DM bezeichnet.
}$^,$\footnote{%
In der Literatur\index{Literatur} und in Filmen\index{Filme} werden nicht
nur klassische Verfahren
benutzt (wie z.~B. die geheime Nachricht, die Sherlock Holmes in der
Geschichte "`Die tanzenden M"annchen"' von Arthur Conan Doyle \index{Arthur 
Conan Doyle} l"ost), sondern auch moderne Verfahren: z.~B. in
% hier die items nicht einr"ucken!
\begin{list}{\textbullet}{\leftmargin10pt}
\item dem Film "`Das Kartenhaus"', 1992, wo sich die Autisten mit Hilfe von
  5- und 6-stelligen Primzahlen unterhalten,
\item der Geschichte "`Der Dialog der Schwestern"' von Dr.~C.~Elsner, 
  \index{Elsner 1999} 1999
  (als PDF dem CrypTool-Paket\index{CrypTool} beigelegt),
  wo die Heldinnen sich vertraulich mit einer Variante des RSA-Verfahrens
  unterhalten.
\item dem Roman "`Das Museum der gestohlenen Erinnerungen"' von Ralf Isau, 
  \index{Isau 2003} Thienemann-Verlag 2003, wo der letzte Teil des Spruches
  nur durch die vereinten Kr"afte der Computergemeinschaft gel"ost werden kann.
\end{list}
}
-Verfahren \glqq beweisen''.

@@

@@

% \newpage  be_24.5.03 raus, da sonst vorherige Seite fast leer.
TABLE 4.1\footnote{%
Die Zahlen dieser Tabelle k"onnen Sie mit Copy und Paste weiter bearbeiten.
}: RSA-Ciphertext

>>>>

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Gruppen und modulare Arithmetik "uber $\mathbb{Z}_n $ und $\mathbb{Z}^*_n $}
\index{Gruppen}
In der Zahlentheorie und in der Kryptographie spielen mathematische 
\glqq {\em Gruppen}'' eine entscheidende Rolle. Von Gruppen spricht man nur, wenn f"ur
eine definierte Menge und eine definierte Relation (eine Operation wie
Addition oder Multiplikation) die folgenden Eigenschaften erf"ullt sind:
\begin{itemize}
\item Abgeschlossenheit \index{Abgeschlossenheit}
\item Existenz des neutralen Elements
\item Existenz eines inversen Elements zu jedem Element und
\item G"ultigkeit des Assoziativgesetzes.
\end{itemize}
Die abgek"urzte mathematische Schreibweise lautet: $(G, +)$ oder $(G,*)$.
\begin{definition}\label{def-zth-zn}
$\mathbb{Z}_n$ :\index{Z@$\mathbb{Z}_n$}
$$\mathbb{Z}_n \text{ umfasst alle ganzen Zahlen von } 0 \text{ bis } n-1: \mathbb{Z}_n = \{0, 1, 2,\cdots, n-2, n-1\}.$$
\end{definition}
$\mathbb{Z}_n$ ist eine h"aufig verwendete endliche Gruppe aus den nat"urlichen Zahlen. Sie 
wird manchmal auch als Restemenge $R$ modulo $n$ bezeichnet.

@@

{\bf Beispiel 2:} \\
Die folgende Tabelle\footnote{%
In \hyperlink{AppArith3b}{Anhang D zu diesem Kapitel} finden Sie den Quelltext zur Berechnung der 
Tabelle mit Mathematica und Pari-GP.
} zeigt die Werte $a^i{\rm ~mod~}45$
f"ur die Exponenten $i = 1, 2, \cdots, 12$ und f"ur die Basen $a = 1, 2, \cdots, 12$
sowie den sich f"ur jedes $a$ daraus ergebenden Wert $ord_{45}(a)$.

@@

\newpage
{\bf Beispiel 4:} \\
Die folgende Tabelle\footnote{%
In \hyperlink{AppArith3c}{Anhang D zu diesem Kapitel} finden Sie den Quelltext zur Berechnung der Tabelle mit Mathematica und Pari-GP.
} beantwortet die Frage, ob die Zahl $7$ eine Primitivwurzel von $46$ ist.
Die Voraussetzung/Bedingung ggT$(7,46)=1$ ist erf"ullt.

@@

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Beweis des RSA-Verfahrens mit Euler-Fermat}\label{rsabeweis}
\hypertarget{RSABeweis}{} Mit dem Satz von Euler-Fermat kann man in der Gruppe $\mathbb{Z}_n^*$ das 
RSA\footnote{%
Das RSA-Verfahren\index{RSA} ist das verbreitetste asymmetrische \index{Verschl""usselung!asymmetrisch} Kryptoverfahren. Es
wurde 1978 von Ronald Rivest, Adi Shamir und Leonard Adleman entwickelt und
kann sowohl zum Signieren wie zum Verschl"usseln eingesetzt werden.
Kryptographen assoziieren mit der Abk"urzung \glqq RSA'' immer dieses
Verfahren $-$ die folgende Anmerkung soll eher humorvoll zeigen, dass man jede
Buchstabenkombination mehrfach belegen kann: in Deutschland gibt es sehr
Interessen-lastige Diskussionen im Gesundheitswesen. Dabei wird mit \glqq RSA''
der vom Gesundheitsministerium geregelte \glqq {\bf R}isiko{\bf S}truktur{\bf A}usgleich'' in der
gesetzlichen Krankenversicherung mit einem j"ahrlichen Volumen von "uber 24
Milliarden DM bezeichnet.
}$^,$\footnote{%
In der Literatur\index{Literatur} und in Filmen\index{Filme} werden nicht
nur klassische Verfahren
benutzt (wie z.~B. die geheime Nachricht, die Sherlock Holmes in der
Geschichte "`Die tanzenden M"annchen"' von Arthur Conan Doyle \index{Arthur 
Conan Doyle} l"ost), sondern auch moderne Verfahren: z.~B. in
% hier die items nicht einr"ucken!
\begin{list}{\textbullet}{\leftmargin10pt}
\item dem Film "`Das Kartenhaus"', 1992, wo sich die Autisten mit Hilfe von
  5- und 6-stelligen Primzahlen unterhalten,
\item der Geschichte "`Der Dialog der Schwestern"' von Dr.~C.~Elsner, 
  \index{Elsner 1999} 1999
  (als PDF dem CrypTool-Paket\index{CrypTool} beigelegt),
  wo die Heldinnen sich vertraulich mit einer Variante des RSA-Verfahrens
  unterhalten,
\item dem Roman "`Das Museum der gestohlenen Erinnerungen"' von Ralf Isau, 
  \index{Isau 2003} Thienemann-Verlag 2003, wo der letzte Teil des Spruches
  nur durch die vereinten Kr"afte der Computergemeinschaft gel"ost werden kann.
\end{list}
}
-Verfahren \glqq beweisen''.

@@

Die 160-stellige Dezimalzahl stammt von der alten Challenge-Liste von RSADSI.
Diese wurde nach der Faktorisierung von RSA-155 (RSA512) zur"uckgezogen,
und es blieben darin noch RSA576, RSA640 etc. (in 64-er Schritten).
Die Primfaktoren von RSA-160 waren aber nicht bekannt.
Deshalb ist dieser Rekord von Prof.\ Frankes Team immer noch die L"osung 
einer alten Challenge, f"ur die es aber von RSADSI kein Geld gibt.

@@

% \newpage  be_24.5.03 raus, da sonst vorherige Seite fast leer.
\vspace{12pt}
TABLE 4.1\footnote{%
Die Zahlen dieser Tabelle k"onnen Sie mit Copy und Paste weiter bearbeiten.
}: RSA-Ciphertext

### 2003-05-30 script/en/chapters/digitalsignatures.tex
fixes for beta6

The document is sent to the recipient together with the signature. The recipient
can then use the sender's public key, the document and the signature to
establish whether or not the signature is correct. In practice, however, the
procedure we have just described has a decisive disadvantage. The signature is
approximately as long as the document itself. To prevent an unnecessary increase
in data traffic, and also for reasons of performance\index{Performance}, we use 
a cryptographic hash function\footnote{%
%xxxxxxxxxx Menüeinträgsbezeichnungen ok?
Hash functions\index{Hash function} are implemented within CrypTool\index{CrypTool} v1.3.04 at
several places. \\
Using menu {\bf Indiv. Procedures \textbackslash{} Hash} you have the possibilities
% hier die items nicht einrücken!
\begin{list}{\textbullet}{\leftmargin10pt\addtolength{\itemsep}{-1.0\baselineskip}}
%\begin{itemize}\addtolength{\itemsep}{-1.0\baselineskip} % um den Abstand zu verringern bebe
\item to apply one of 6 hash functions to the content of the current window, \\
\item to calculate the hash value of a file, \\
\item to test, how changes to a text change the according hash value and \\
\item to perform a simulation, how digital signatures could be attacked 
      by a targeted search for hash value collisions\index{collisions}.
\end{list}
}.

% --------------------------------------------------------------------------
\subsection{Hash functions}
\index{Hash function}
A cryptographic \index{Hash function} {\em hash function}\footnote{%
Secure hash algorithms compute a condensed representation of 
electronic data (message).
When a message is input to a hash algorithm, the result is an output called
a message digest. The message digests typically range in length 
from 128 to 512 bits, depending on the algorithm. 
Secure hash algorithms are typically used with
other cryptographic algorithms, such as digital signature algorithms and
keyed-hash message authentication codes, or in the generation of random
numbers (bits).
}
maps a message of any
length to a string of characters with a constant size,
the \index{Hash value} hash value. It should be practically impossible, for a
given number, to find a message that has precisely this number as hash value.
Furthermore, it should be practically impossible to find two messages with the
same hash value\index{collisions}. 
In both cases the final signature procedure would include weaknesses.

So far, no formal proof has been found that perfectly secure cryptographic
hash functions exist. However, there are several good candidates that have
not yet shown any weak points in their structure in practice 
(e.g. \index{SHA-1} SHA--1\footnote{%
SHA-1 \index{SHA-1} is a 160 bit hashfunction specified in FIPS 180-1, 
ANSI X9.30 Part 2 and \cite{FIPS186}.\\
SHA means Secure Hash Algorithm, and is widely used, 
e.g. with DSA, RSA or ECDSA.\\
The current standard \cite{FIPS180-2} defines four secure hash algorithms
-- SHA-1, SHA-256, SHA-384, and SHA-512.

For these hash algorithms there are also validation tests defined in the
test suite FIPS 140-2.

@@

}  or \index{RIPEMD-160} RIPEMD-­160\footnote{%
RIPEMD-160, RIPEMD-128 and the optional extension RIPEMD-256 have object
identifiers defined by the ISO-identified organization TeleTrusT, both 
as hash algorithm and in combination with RSA.

RIPEMD-160 is also part of the ISO/IEC international standard 
ISO/IEC 10118-3:1998 on dedicated hash functions, together with 
RIPEMD-128 and SHA-1.

Further details: \\
http://www.esat.kuleuven.ac.be/~bosselae/ripemd160.html \\
http://www.ietf.org/rfc/rfc2857.txt ("The Use of HMAC-RIPEMD-160-96 within ESP and AH").
}
).

>>>>

The document is sent to the recipient together with the signature. The recipient
can then use the sender's public key, the document and the signature to
establish whether or not the signature is correct. In practice, however, the
procedure we have just described has a decisive disadvantage. The signature is
approximately as long as the document itself. To prevent an unnecessary increase
in data traffic, and also for reasons of performance\index{Performance}, we use 
a cryptographic hash function\footnote{%
%xxxxxxxxxx Menüeinträgsbezeichnungen ok?
Hash functions\index{Hash function} are implemented within CrypTool\index{CrypTool} v1.3.04 at
several places. \\
Using menu {\bf Indiv. Procedures \textbackslash{} Hash} you have the possibilities
% hier die items nicht einrücken!
\begin{list}{\textbullet}{\leftmargin10pt\addtolength{\itemsep}{-1.0\baselineskip}}
%\begin{itemize}\addtolength{\itemsep}{-1.0\baselineskip} % um den Abstand zu verringern bebe
\item to apply one of 6 hash functions to the content of the current window, \\
\item to calculate the hash value of a file, \\
\item to test, how changes to a text change the according hash value and \\
\item to perform a simulation, how digital signatures could be attacked 
      by a targeted search for hash value collisions\index{Collision}.
\end{list}
}.

% --------------------------------------------------------------------------
\subsection{Hash functions}
\index{Hash function}
A cryptographic \index{Hash function} {\em hash function}\footnote{%
Secure hash algorithms compute a condensed representation of 
electronic data (message).
When a message is input to a hash algorithm, the result is an output called
a message digest. The message digests typically range in length 
from 128 to 512 bits, depending on the algorithm. 
Secure hash algorithms are typically used with
other cryptographic algorithms, such as digital signature algorithms and
keyed-hash message authentication codes, or in the generation of random
numbers (bits).
}
maps a message of any
length to a string of characters with a constant size,
the \index{Hash value} hash value. It should be practically impossible, for a
given number, to find a message that has precisely this number as hash value.
Furthermore, it should be practically impossible to find two messages with the
same hash value\index{collisions}. 
In both cases the final signature procedure would have weaknesses.

So far, no formal proof has been found that perfectly secure cryptographic
hash functions exist. However, there are several good candidates that have
not yet shown any weak points in their structure in practice 
(e.g. \index{SHA-1} SHA--1\footnote{%
SHA-1 \index{SHA-1} is a 160 bit hash function specified in FIPS 180-1, 
ANSI X9.30 Part 2 and \cite{FIPS186}.\\
SHA means Secure Hash Algorithm, and is widely used, 
e.g. with DSA, RSA or ECDSA.\\
The current standard \cite{FIPS180-2} defines four secure hash algorithms
-- SHA-1, SHA-256, SHA-384, and SHA-512.
For these hash algorithms there are also validation tests defined in the
test suite FIPS 140-2.

@@

}  or \index{RIPEMD-160} RIPEMD-­160\footnote{%
RIPEMD-160, RIPEMD-128 and the optional extension RIPEMD-256 have object
identifiers defined by the ISO-identified organization TeleTrusT, both 
as hash algorithm and in combination with RSA.
RIPEMD-160 is also part of the ISO/IEC international standard 
ISO/IEC 10118-3:1998 on dedicated hash functions, together with 
RIPEMD-128 and SHA-1.

Further details: \\
http://www.esat.kuleuven.ac.be/~bosselae/ripemd160.html \\
http://www.ietf.org/rfc/rfc2857.txt (``The Use of HMAC-RIPEMD-160-96 within ESP and AH'').
}).

### 2003-05-30 script/en/chapters/ellipticcurves.tex
*** empty log message ***
fixes for beta6

\newpage
\section{Elliptic Curves}
\index{Elliptic curves}
\hypertarget{ellcurve}{}
(Filipovics B. / B\"uger M. / Esslinger B. / Oyono R., April 2000, Updates: Dec. 2001, June 2002, Mar. 2003)

@@

Enough about the role of elliptic curves in pure mathematics. In 1985 Neal
Koblitz and Victor Miller independently suggested using elliptic curves in
cryptography. Elliptic curves have thus also found a concrete practical
application. Another interesting field of application for elliptic curves is for
factorising whole numbers. (For example the RSA cryptography system is based on
the \index{Complexity} difficulty/complexity of finding prime factors of an
extremely large number.) In this area, procedures based on elliptic curves have
been investigated and partially used since 1987 (a study by H.W. Lenstra). There
are also prime number tests\index{Prime number!test} based on elliptic curves.

Elliptic curves are used differently in the various areas. Encryption procedures
based on elliptic curves are based on the difficulty of a problem known as
elliptic curve discrete logarithm\index{Logarithm problem!discrete}. The factorisation of whole numbers uses the
fact that a large number of elliptic curves can be generated for a natural
composite number $n$ with several prime factors; however, these curves are not
then groups for composite $n$. \hyperlink{faktell}{More information about this
can be found under Factorisation using elliptic curves.}

@@

\subsubsection{Factorisation using elliptic curves}

\hypertarget{faktell}{}

>>>>

\newpage
\section{Elliptic Curves}
\index{Elliptic curve}
\hypertarget{ellcurve}{}
(Filipovics B. / B\"uger M. / Esslinger B. / Oyono R., April 2000, Updates: Dec. 2001, June 2002, Mar. 2003)

@@

Enough about the role of elliptic curves in pure mathematics. In 1985 Neal
Koblitz\index{Koblitz Neal} and Victor Miller\index{Miller Victor} independently suggested using elliptic curves in
cryptography. Elliptic curves have thus also found a concrete practical
application. Another interesting field of application for elliptic curves is for
factorising whole numbers. (For example the RSA cryptography system is based on
the \index{Complexity} difficulty/complexity of finding prime factors of an
extremely large number.) In this area, procedures based on elliptic curves have
been investigated and partially used since 1987 (a study by H.W. Lenstra). There
are also prime number tests\index{Prime number!test} based on elliptic curves.

Elliptic curves are used differently in the various areas. Encryption procedures
based on elliptic curves are based on the difficulty of a problem known as
elliptic curve discrete logarithm\index{Logarithm problem!discrete}. The factorisation of whole numbers uses the
fact that a large number of elliptic curves can be generated for a natural
composite number $n$ with several prime factors; however, these curves are not
then groups for composite $n$. More information about this can be found in
section \ref{faktell}.

@@

\subsubsection{Factorisation using elliptic curves}
\label{faktell}

### 2003-05-30 script/en/chapters/menus.tex
fixes for beta6

\pagebreak
\enlargethispage{1cm}
\section{CrypTool Menus}
\hypertarget{appendix-menutree}{}\label{s:appendix-menutree}
This appendix contains the complete menu tree of CrypTool. 
Please notice that not all menu items are always available, because the
menu tree is built up dynamically: which menu items are available depends
on the type of the currently active document window.
The brute-force analysis for DES e.~g. is only available, if the active
window is opened in the hexadecimal view. On the other hand the menu item
``Generate Random Numbers\dots'' is always available. 
Within the figures \ref{menu-detail-1} and \ref{menu-detail-2} 
this relation is shown in the brackets after the menu items,
e.~g. ``Find\dots[ATH]'' means, that the function ``Find\dots'' is only
available for ASC-, text- and hexadecimal windows.
\begin{center}
\begin{tabular}{rl}
\bf Code letter & \bf Type of document \\
M & (no document opened)\\
A & ASC\\
T & Text\\
H & Hexadecimal\\
P & Plot\\
\end{tabular}
\end{center}

@@

\begin{figure}[b]
\begin{center}
%\frame{
\includegraphics[scale=1.1, clip, viewport=404 117 828 598]{figures/cryptool-menu-detail-en}
%}
\caption{Detailed view of the CrypTool menu tree -- part 2}
\label{menu-detail-2}
\end{center}
\end{figure}
\clearpage
\begin{figure}[hb]
\begin{center}
\vspace{-30pt}
%\frame{
\includegraphics[scale=0.75, angle=270, viewport=14 5 829 590]{figures/cryptool-menu-en}
%}
\caption{Complete overview of the CrypTool menu tree} 
\label{menuoverview}
\end{center}
\end{figure}
\clearpage

>>>>

\pagebreak
\enlargethispage{1cm}
\section{CrypTool Menus}
\hypertarget{appendix-menutree}{}\label{s:appendix-menutree}
This appendix contains the complete menu tree of CrypTool version 1.3.04. 
Please notice that not all menu items are always available, because the
menu tree is built up dynamically: which menu items are available depends
on the type of the currently active document window.
The brute-force analysis for DES e.~g. is only available, if the active
window is opened in the hexadecimal view. On the other hand the menu item
``Generate Random Numbers\dots'' is always available. 
Within the figures \ref{menu-detail-1} and \ref{menu-detail-2} 
this relation is shown in the brackets after the menu items,
e.~g. ``Find\dots[ATH]'' means, that the function ``Find\dots'' is only
available for ASC-, text- and hexadecimal windows.
\begin{center}
\begin{tabular}{rl}
\bf Code letter & \bf Type of document \\
M & (no document opened)\\
A & ASC\\
T & Text\\
H & Hexadecimal\\
P & Plot\\
\end{tabular}
\end{center}

@@

\begin{figure}[b]
\begin{center}
%\frame{
\includegraphics[scale=1.1, clip, viewport=404 117 831 598]{figures/cryptool-menu-detail-en}
%}
\caption{Detailed view of the CrypTool menu tree -- part 2}
\label{menu-detail-2}
\end{center}
\end{figure}
\clearpage
\begin{figure}[hb]
\begin{center}
\vspace{-30pt}
%\frame{
\includegraphics[scale=0.75, angle=270, viewport=14 5 829 590]{figures/cryptool-menu-en}
%}
\caption{Complete overview of the CrypTool menu tree} 
\label{menuoverview}
\end{center}
\end{figure}
\clearpage

### 2003-05-30 script/en/chapters/moderncryptography.tex
fixes for beta6

\newpage
%**************************************************************************************************************************
\section{The Mathematical Ideas behind Modern Cryptography}
(Oyono R. / Esslinger B., Sep. 2000, Updates Nov. 2000, Feb. 2003)
\subsection{One way functions with trapdoor and complexity classes}
\index{Cryptography!modern} \index{One way function} \hypertarget{OneWayFunktion1}{}
A {\bf one way function} is a (one-to-one) function that can be calculated 
efficiently, but whose inverse is extremely complicated and practically 
impossible to calculate.\par

>>>>

\newpage
%**************************************************************************************************************************
\section{The Mathematical Ideas behind Modern Cryptography}
(Oyono R. / Esslinger B., Sep. 2000, Updates Nov. 2000, Feb. 2003)
\subsection{One way functions with trapdoor and complexity classes}
\index{Cryptography!modern} \index{One way function} \hypertarget{OneWayFunktion1}{}
A {\bf one way function} is a function that can be calculated 
efficiently, but whose inverse is extremely complicated and practically 
impossible to calculate.\par
